---
title: "R Notebook"
output:
  word_document: default
  pdf_document: default
  html_document:
    df_print: paged
---

# Sample of UFC Data analysis

```{r}
#Read Data
UFCdata <- read.csv("data.csv", header = TRUE)
head(UFCdata)
```

# Clean Data
```{r}
#Look at structure of Data Set
str(UFCdata)

##Do not want to use these columns (Referee, Location, Title, Rounds) for our prediction, want to focus on features aligned to the fighter (specifically we want to focus on physical features for our model). We will also remove the date column later.
UFCdata <- UFCdata[c(-3,-5,-7,-9)]
```

```{r}
# R_fighter & B_fighter are the fighter's name which will be the inputs of the model. They will not be used for prediction. They are essentially used as labels for whoever wins the fight. 
# Will have to extract names from data later
```

```{r}
# Winner currently has three levels; however the model we will use will either predicting a "Win" or "Loss". Therefore, we will remove observations that end in a draw. The number is relatively small, therefore is not concerning, but is limitation for our future model.
sum(UFCdata$Winner == "Draw")
UFCdata <- UFCdata[UFCdata$Winner !="Draw",]
UFCdata$Winner <- factor(UFCdata$Winner)
levels(UFCdata$Winner)
```


```{r}
# Check for NAs 
sapply(UFCdata, function(x) {sum(is.na(x))})

# The NAs in R_age, B_age, R_reach_cms, & B_reach_cms pose a problem as we believe these physical features to have an effect on the fight outcome and simply appear to be missing. Thankfully, these points make up a relatively small part of the data set, therefore observations in these columns will be removed
UFCdata <- UFCdata[is.na(UFCdata$R_age) == FALSE,]
UFCdata <- UFCdata[is.na(UFCdata$B_age) == FALSE,]
UFCdata <- UFCdata[is.na(UFCdata$R_Reach_cms) == FALSE,]
UFCdata <- UFCdata[is.na(UFCdata$B_Reach_cms) == FALSE,]

# First, there are a lot of NAs remaining in the data. However, the NAs that appear are in features relating to UFC career stats. These NAs mean that it’s the fighters first fight, therefore do not have data from pervious fights to fill out the columns. These cells will be converted to 0s showing the fighters lack of experience and true stats at the moment. It is important to keep these observations as it will help us account for first time fighters.
UFCdata[is.na(UFCdata)] <- 0
```

```{r}
# Double Check Data
sapply(UFCdata, function(x) {sum(is.na(x))})
```

```{r}
#Check for Missing Data
sapply(UFCdata, function(x){sum(x=="")})

# It appears a few observations in B_Stance & R_Stance are missing. These observations will be removed. 
UFCdata <- UFCdata[UFCdata$B_Stance !="",]
UFCdata <- UFCdata[UFCdata$R_Stance !="",]
```

```{r}
#Double Check Data
sapply(UFCdata, function(x){sum(x=="")})
```

```{r}
# Due to the rules of UFC drastically changing after 2000, the data set should only contain observations past 2000.
# First, Change structure of date from Factor to Date
str(UFCdata$date)
UFCdata$date <- as.Date(UFCdata$date)

# Split Data Set Between Old UFC & Modern UFC
UFCdata <- UFCdata[UFCdata$date > "1999-12-31",]

# Now remove the date feature for the same reasons described in the second chunk.
UFCdata <- UFCdata[c(-3)]
```

```{r}
# Different subsets of the data will be created based on the weight class. We are doing this because fighters from two different weight classes would never fight each other under UFC rules. We will be conducting analysis on different weight classes to produce a more accurate model for each class
```

```{r}
# Now lets subset the data into different weight classes
# See Number of Weight Classes
levels(UFCdata$weight_class)
```

```{r}
#Create Datasets for Different Weight Classes
weightclass <- levels(UFCdata$weight_class)
for(i in 1:14) {
  name <- weightclass[i]
  assign(name, UFCdata[UFCdata$weight_class == weightclass[i],])
}
```

# Exploratory 

```{r}
#Distribution between Weight Classes
barplot((summary(UFCdata$weight_class)),col="pink")
```
Number of observations vary between weight classes, due to some class being more popular than others.


```{r}
# Distribution of matches by number of rounds
UFCdataRounds <- read.csv("data.csv", header = TRUE)
UFCdataRounds$no_of_rounds <- as.factor(as.character(UFCdataRounds$no_))
barplot((summary(UFCdataRounds$no_of_rounds)))
```
There is only a small number of 5 round fights which are typically title fights, these were remove because we want our model to focus on physical attributes. 


```{r}
# Load ggplot2
library(ggplot2)
```


```{r}
#tittle 
data <- data.frame(
  group=c("False","True"),
  value=c(.934,.065)
)

ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void()
```

```{r}
# Distribution of age in the UFC (Red Fighters)
UFCdata$R_age <- as.factor(as.character(UFCdata$R_age))
barplot((summary(UFCdata$R_age)), col = "red")
```

```{r}
# Distribution of age in the UFC (Blue Fighters)
UFCdata$B_age <- as.factor(as.character(UFCdata$B_age))
barplot((summary(UFCdata$B_age)),col = "lightblue")
```

```{r}
# Histograms of fighters accuracy
UFCdata$B_body_accuracy <- UFCdata$B_avg_BODY_landed/UFCdata$B_avg_BODY_att
hist(UFCdata$B_body_accuracy)
UFCdata$R_body_accuracy <- UFCdata$R_avg_BODY_landed/UFCdata$R_avg_BODY_att
hist(UFCdata$R_body_accuracy)
UFCdata$body_accuracy_diff <- UFCdata$R_body_accuracy - UFCdata$B_body_accuracy
hist(UFCdata$body_accuracy_diff)
```
With this data we can calculate cool fighter stats like efficiency. 


```{r}
# Number of fights throughout the years
UFCdataRounds$date <- as.character(UFCdataRounds$date)
UFCdataRounds$year <- as.factor(substring(UFCdataRounds$date,1,4))
barplot(summary(UFCdataRounds$year))
```
The number of observations grows as the years increases, due to UFC becoming more popular. 

```{r}
#Breakdown of Red Winners, Blue Winners, Draws
UFCdata$Winner <- as.factor(UFCdata$Winner)

data <- data.frame(
  group=c("Red Wins","Blue Wins","Draw"),
  value=c(.675,.309,.016)
)

ggplot(data, aes(x="", y=value, fill=group)) +
  geom_bar(stat="identity", width=1, color="white") +
  coord_polar("y", start=0) +
  
  theme_void()
```
There is major data imbalance that our later models will try to fix. THE CORNER SHOULD NOT EFFECT A FIGHTERS CHANCE OF WINNING, the reason there is this data imbalance is UFC promoters have put their champions and favourites in that corner, causing that side to win more. However the model we make should ignore the corner they are standing in as it should not affect the result.

# Max & Min Weight of (Heavyweight) Class
```{r}
MinFighterWeights <- c(min(Heavyweight$B_Weight_lbs),min(Heavyweight$R_Weight_lbs))
min(MinFighterWeights)

MaxFighterWeights <- c(max(Heavyweight$B_Weight_lbs),max(Heavyweight$R_Weight_lbs))
max(MaxFighterWeights)

```

# Average Height of (Heavyweight) Class
```{r}
AverageHeights <- c(mean(Heavyweight$B_Height_cms),mean(Heavyweight$B_Height_cms))
mean(AverageHeights) 
```

# Summary Statistics of (Heavyweight) Class
```{r}
# NOT running the code because it is extremely long
"summary(Heavyweight)
summary(Heavyweight$B_age)
summary(Heavyweight$B_Height_cms)
summary(Heavyweight$B_Weight_lbs)"
```


#-------------------------------------------------------------------------------------------------------------
# MAIN QUESTION WE WANT TO ANSWER: Who is likely to win a UFC fight based on Physcial Attributes?
The hope of this model is that a potential fighter entering the UFC can use model to see what weight class they would be most successful in based on their physical attributes. Therefore, a fighter can prioritize and train for a weight class that he would have an advantage in due to his physical attributes. HOWEVER, we are aware a physical attributes should not fully explain the result of the fight. The sport usually requires experience, training, and mental strategy to be successful, but physical attributes can give some advantages to fighters.
#--------------------------------------------------------------------------------------------------------------

# -------------------------------------------------------------------------------------------------------------
 
# Who is likely to win a UFC fight based on Physcial Attributes

#--------------------------------------------------------------------------------------------------------------

#------------------------------------------------
#Welterweight
#------------------------------------------------

Read the data
```{r}
#Select Physical Features
WelterweightPhy <-  Welterweight[c(1,2,3,69,70,71,136,137,138,139,140)]
head(WelterweightPhy)
```

# Create testing & training sets for SVM, K-NN, & Neural Networks

```{r}
# Normalize All the Attributes
for (i in 4:dim(WelterweightPhy)[2])
{
  WelterweightPhy[i] = (WelterweightPhy[i]-mean(WelterweightPhy[,i]))/sd(WelterweightPhy[,i])
}

# Display the first few elements of the standarized data
head(WelterweightPhy)
```

```{r}
# Create Training and Testing Sets
num_samples = dim(WelterweightPhy)[1]
sampling.rate = 0.80
training <- sample(1:num_samples, sampling.rate * num_samples, replace=FALSE)
trainingSet <- WelterweightPhy[training, ]
testing <- setdiff(1:num_samples,training)
testingSet <- WelterweightPhy[testing, ]
```

```{r}
# Slpit Fighter's names from sets
trainingSet2 <- trainingSet[c(-1,-2)]
testingSet2 <- testingSet[c(-1,-2)]
```


# K-Nearest Neighbors

```{r}
# Get the features of the training set
trainingfeatures <- subset(trainingSet2, select=c(-Winner))
# Get the labels of the training set
traininglabels <- trainingSet2$Winner
# Get the features of the testing set
testingfeatures <- subset(testingSet2, select=c(-Winner))
```

```{r}
# Load the classification library
library(class)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(1,50, 1))
{
  predictedLabels = knn(trainingfeatures,testingfeatures,traininglabels,k=i)
sizeTestSet = dim(testingSet2)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```


```{r}
# Create Predictions Table to store predictions from different models
PredictionsTable <- data.frame(matrix(ncol = 3, nrow = dim(testingSet)[1]))
x <- c("KNN","SVM","NN")
colnames(PredictionsTable) <- x
# Fill in KNN Col
PredictionsTable$KNN <- predictedLabels
```


```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```



#SVM

```{r}
# Load the SVM Library
library(e1071)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(10,100, 10))
{
  svmModel <- svm(Winner~., data=trainingSet2, kernel="linear", cost=i)
  predictedLabels <-predict(svmModel, testingSet2)
sizeTestSet = dim(testingSet)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```

```{r}
# Fill in Table
PredictionsTable$SVM <- predictedLabels
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


# Neural Networks



```{r}
# Convert the characters to numeric values
# training
trainingSet2$Winner <- as.character(trainingSet2$Winner)
trainingSet2$Winner[trainingSet2$Winner=="Red"] = 0
trainingSet2$Winner[trainingSet2$Winner=="Blue"] = 1
trainingSet2$Winner <- as.numeric(trainingSet2$Winner)
# testing
testingSet2$Winner <- as.character(testingSet2$Winner)
testingSet2$Winner[testingSet2$Winner=="Red"] = 0
testingSet2$Winner[testingSet2$Winner=="Blue"] = 1
testingSet2$Winner <- as.numeric(testingSet2$Winner)
```

```{r}
#Load the Neural Net Library
library(neuralnet)
#Train a Neural Network. Note that the formula should be explicitly specified.
#Since we are doing a classification, we need to set linear.output=FALSE
#We will also use a Neural network with 2 hidden layers, with 7, 5 neurons respectively
nnModel <- neuralnet(Winner~., data=trainingSet2, hidden=c(7,5), linear.output=FALSE)
```

```{r}
# Plot
plot(nnModel)
```

```{r}
# Perform prdictions for the testing set
predictedLabels <-neuralnet::compute(nnModel, testingSet2[,2:9])

#The result is stored in predictedLabels$net.result
#Note that we need to round the predictions to get 0 or 1
predictedLabels<-round(predictedLabels$net.result)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(predictedLabels != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 0)
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 1)
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Fill in Table
PredictionsTable$NN <- predictedLabels
```

```{r}
# Convert #s to labels
PredictionsTable$NN <- as.character(PredictionsTable$NN)
PredictionsTable$NN[PredictionsTable$NN== "0" ] = "Red"
PredictionsTable$NN[PredictionsTable$NN== "1"] = "Blue"
```


# Convert Back
```{r}
# Convert back
# training
trainingSet2$Winner[trainingSet2$Winner== 0 ] = "Red"
trainingSet2$Winner[trainingSet2$Winner== 1 ] = "Blue"
trainingSet2$Winner <- as.factor(trainingSet2$Winner)
# testing
testingSet2$Winner[testingSet2$Winner== 0 ] = "Red"
testingSet2$Winner[testingSet2$Winner== 1 ] = "Blue"
testingSet2$Winner <- as.factor(testingSet2$Winner)
```




# Fixing class Imbalance 
Most the data has the red class winning, which should not be the case due to qualitative reasons (a lot to explain). Both red and blue class should equally represented as possible. 

Ideally would convert winning red observations into winning blue observations, but do not know who to do that yet.

Instead, I will try to achieve this by over and under sampling. 

```{r}
library(ROSE)
```

```{r}
#See # of wins between colors
table(trainingSet$Winner)
```

```{r}
ROSE <- ovun.sample(Winner~.,data = trainingSet2, method = "both", p = 0.5, seed = 111, N = 645)$data
table(ROSE$Winner)
```



# K-NN

```{r}
# Get the features of the training set
trainingfeatures <- subset(ROSE, select=c(-Winner))
# Get the labels of the training set
traininglabels <- ROSE$Winner
# Get the features of the testing set
testingfeatures <- subset(testingSet2, select=c(-Winner))
```

```{r}
# Load the classification library
library(class)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(1,50, 1))
{
  predictedLabels = knn(trainingfeatures,testingfeatures,traininglabels,k=i)
sizeTestSet = dim(testingSet2)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Fill in Table
PredictionsTable$KNN_RB <- predictedLabels
```


#SVM Model

```{r}
# Load the SVM Library
library(e1071)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(10,100, 10))
{
  svmModel <- svm(Winner~., data=ROSE, kernel="linear", cost=i)
  predictedLabels <-predict(svmModel, testingSet2)
sizeTestSet = dim(testingSet)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Fill in Table
PredictionsTable$SVM_RB <- predictedLabels
```

# Combining predictions of models

```{r}
head(PredictionsTable)
```

```{r}
PredictionsTable$KNN <- as.character(PredictionsTable$KNN)
PredictionsTable$SVM <- as.character(PredictionsTable$SVM)
PredictionsTable$KNN_RB <- as.character(PredictionsTable$KNN_RB)
PredictionsTable$SVM_RB <- as.character(PredictionsTable$SVM_RB)
```


```{r}
PredictionsTable[PredictionsTable == "Red"] <- 1
PredictionsTable[PredictionsTable == "Blue"] <- 0
```

```{r}
PredictionsTable$KNN <- as.numeric(PredictionsTable$KNN)
PredictionsTable$SVM <- as.numeric(PredictionsTable$SVM)
PredictionsTable$NN <- as.numeric(PredictionsTable$NN)
PredictionsTable$KNN_RB <- as.numeric(PredictionsTable$KNN_RB)
PredictionsTable$SVM_RB <- as.numeric(PredictionsTable$SVM_RB)
```

```{r}
#Scoring 
PredictionsTable$Score <- rowSums(PredictionsTable)
```

```{r}
# Prediction 
PredictionsTable$Winner <- ifelse(PredictionsTable$Score > 2.5, "Red", "Blue")
```

```{r}
#Input Winner's Name
PredictionsTable$Red <- testingSet$R_fighter
PredictionsTable$Blue <- testingSet$B_fighter
PredictionsTable$Winner_Name <- ifelse(PredictionsTable$Winner == "Red", as.character(PredictionsTable$Red), as.character(PredictionsTable$Blue))
```

# Results

```{r}
ResultsW <- PredictionsTable
head(ResultsW)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(PredictionsTable$Winner != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (PredictionsTable$Winner == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (PredictionsTable$Winner == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


#--------------------------------------------------------------------------------------------------------------
# Another way to fix the data imbalance is to "flip" the data from Red to Blue
#--------------------------------------------------------------------------------------------------------------

#--------------------------------------------------------------------------------------------------------------
#Welterweight
#--------------------------------------------------------------------------------------------------------------

Read the data
```{r}
#Select Physical Features
WelterweightPhy <-  Welterweight[c(1,2,3,69,70,71,136,137,138,139,140)]
head(WelterweightPhy)
```

```{r}
WelterweightPhyFlip = WelterweightPhy[,c(3,2,4:6,10,1,7:9,11)]
WelterweightPhyFlip$B_fighter <- as.character(WelterweightPhyFlip$B_fighter)
WelterweightPhyFlip$R_fighter <- as.character(WelterweightPhyFlip$R_fighter)
head(WelterweightPhyFlip)
```

```{r}
for(i in 1:dim(WelterweightPhyFlip)[1]){
  if(WelterweightPhyFlip$Winner[i] == "Red"){
  randomNum = runif(1)
    if(randomNum < 0.25){
      tempDf = WelterweightPhyFlip[i,2:6]
      WelterweightPhyFlip[i,2:6] = WelterweightPhyFlip[i,7:11]
      WelterweightPhyFlip[i,7:11] = tempDf[1:5]
      WelterweightPhyFlip[i,1] = "Blue"
    }
  }
}
WelterweightPhyFlip$B_fighter <- as.factor(WelterweightPhyFlip$B_fighter)
WelterweightPhyFlip$R_fighter <- as.factor(WelterweightPhyFlip$R_fighter)
head(WelterweightPhyFlip)
```

```{r}
table(WelterweightPhyFlip$Winner)
WelterweightPhyFlip <- WelterweightPhyFlip[c(2,7,1,3,4,5,6,8,9,10,11)]
```

# Create testing & training sets for SVM, K-NN, & Neural Networks

```{r}
# Normalize All the Attributes
for (i in 4:dim(WelterweightPhyFlip)[2])
{
  WelterweightPhyFlip[i] = (WelterweightPhyFlip[i]-mean(WelterweightPhyFlip[,i]))/sd(WelterweightPhyFlip[,i])
}

# Display the first few elements of the standarized data
head(WelterweightPhyFlip)
```

```{r}
# Create Training and Testing Sets
num_samples = dim(WelterweightPhyFlip)[1]
sampling.rate = 0.80
training <- sample(1:num_samples, sampling.rate * num_samples, replace=FALSE)
trainingSet <- WelterweightPhyFlip[training, ]
testing <- setdiff(1:num_samples,training)
testingSet <- WelterweightPhyFlip[testing, ]
```

```{r}
# Slpit Fighter's names from sets
trainingSet2 <- trainingSet[c(-1,-2)]
testingSet2 <- testingSet[c(-1,-2)]
```

# K-Nearest Neighbors

```{r}
# Get the features of the training set
trainingfeatures <- subset(trainingSet2, select=c(-Winner))
# Get the labels of the training set
traininglabels <- trainingSet2$Winner
# Get the features of the testing set
testingfeatures <- subset(testingSet2, select=c(-Winner))
```

```{r}
# Load the classification library
library(class)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(1,50, 1))
{
  predictedLabels = knn(trainingfeatures,testingfeatures,traininglabels,k=i)
sizeTestSet = dim(testingSet2)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```


```{r}
# Create Predictions Table to store predictions from different models
PredictionsTable <- data.frame(matrix(ncol = 3, nrow = dim(testingSet)[1]))
x <- c("KNN","SVM","NN")
colnames(PredictionsTable) <- x
# Fill in KNN Col
PredictionsTable$KNN <- predictedLabels
```


```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


#SVM

```{r}
# Load the SVM Library
library(e1071)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(10,100, 10))
{
  svmModel <- svm(Winner~., data=trainingSet2, kernel="linear", cost=i)
  predictedLabels <-predict(svmModel, testingSet2)
sizeTestSet = dim(testingSet)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```


```{r}
# Fill in Table
PredictionsTable$SVM <- predictedLabels
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


# Neural Networks

```{r}
# Convert the characters to numeric values
# training
trainingSet2$Winner <- as.character(trainingSet2$Winner)
trainingSet2$Winner[trainingSet2$Winner=="Red"] = 0
trainingSet2$Winner[trainingSet2$Winner=="Blue"] = 1
trainingSet2$Winner <- as.numeric(trainingSet2$Winner)
# testing
testingSet2$Winner <- as.character(testingSet2$Winner)
testingSet2$Winner[testingSet2$Winner=="Red"] = 0
testingSet2$Winner[testingSet2$Winner=="Blue"] = 1
testingSet2$Winner <- as.numeric(testingSet2$Winner)
```

```{r}
#Load the Neural Net Library
library(neuralnet)
#Train a Neural Network. Note that the formula should be explicitly specified.
#Since we are doing a classification, we need to set linear.output=FALSE
#We will also use a Neural network with 2 hidden layers, with 7, 5 neurons respectively
nnModel <- neuralnet(Winner~., data=trainingSet2, hidden=c(7,5), linear.output=FALSE)
```

```{r}
# Plot
plot(nnModel)
```

```{r}
# Perform prdictions for the testing set
predictedLabels <- neuralnet::compute(nnModel, testingSet2[,2:9])

#The result is stored in predictedLabels$net.result
#Note that we need to round the predictions to get 0 or 1
predictedLabels<-round(predictedLabels$net.result)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(predictedLabels != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 0)
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 1)
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```
```{r}
# Fill in Table
PredictionsTable$NN <- predictedLabels
```

```{r}
# Convert #s to labels
PredictionsTable$NN <- as.character(PredictionsTable$NN)
PredictionsTable$NN[PredictionsTable$NN== "0" ] = "Red"
PredictionsTable$NN[PredictionsTable$NN== "1"] = "Blue"
```


# Convert Back
```{r}
# Convert back
# training
trainingSet2$Winner[trainingSet2$Winner== 0 ] = "Red"
trainingSet2$Winner[trainingSet2$Winner== 1 ] = "Blue"
trainingSet2$Winner <- as.factor(trainingSet2$Winner)
# testing
testingSet2$Winner[testingSet2$Winner== 0 ] = "Red"
testingSet2$Winner[testingSet2$Winner== 1 ] = "Blue"
testingSet2$Winner <- as.factor(testingSet2$Winner)
```

# Combining predictions of models

```{r}
head(PredictionsTable)
```

```{r}
PredictionsTable$KNN <- as.character(PredictionsTable$KNN)
PredictionsTable$SVM <- as.character(PredictionsTable$SVM)
```


```{r}
PredictionsTable[PredictionsTable == "Red"] <- 1
PredictionsTable[PredictionsTable == "Blue"] <- 0
```

```{r}
PredictionsTable$KNN <- as.numeric(PredictionsTable$KNN)
PredictionsTable$SVM <- as.numeric(PredictionsTable$SVM)
PredictionsTable$NN <- as.numeric(PredictionsTable$NN)
```

```{r}
#Scoring 
PredictionsTable$Score <- rowSums(PredictionsTable)
```

```{r}
# Prediction 
PredictionsTable$Winner <- ifelse(PredictionsTable$Score > 1.5, "Red", "Blue")
```

```{r}
#Input Winner's Name
PredictionsTable$Red <- testingSet$R_fighter
PredictionsTable$Blue <- testingSet$B_fighter
PredictionsTable$Winner_Name <- ifelse(PredictionsTable$Winner == "Red", as.character(PredictionsTable$Red), as.character(PredictionsTable$Blue))
```

# Results

```{r}
ResultsFlip1 <- PredictionsTable
head(ResultsFlip1)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(PredictionsTable$Winner != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (PredictionsTable$Winner == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (PredictionsTable$Winner == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


#--------------------------------------------------------------------------------------------------------------
#Combining Features 
#--------------------------------------------------------------------------------------------------------------

# Welterweight

Read the data
```{r}
#Select Physical Features
WelterweightPhy <- Welterweight[c(1,2,3,69,70,71,136,137,138,139,140)]
head(WelterweightPhy)
#Take the difference between features 
WelterweightPhy$Height <- WelterweightPhy$B_Height_cms -WelterweightPhy$R_Height_cms
WelterweightPhy$Reach <- WelterweightPhy$B_Reach_cms - WelterweightPhy$R_Reach_cms
WelterweightPhy$Weight <- WelterweightPhy$B_Weight_lbs - WelterweightPhy$R_Weight_lbs
WelterweightPhy$Age <- WelterweightPhy$B_age - WelterweightPhy$R_age
WelterweightPhy$R_H_RF <- WelterweightPhy$R_Reach_cms/WelterweightPhy$R_Height_cms
WelterweightPhy$R_H_BF <- WelterweightPhy$B_Reach_cms/WelterweightPhy$B_Height_cms
WelterweightPhy <- WelterweightPhy[c(1,2,3,12,13,14,15,16,17)]
```
# Create testing & training sets for SVM, K-NN, & Neural Networks

```{r}
# Normalize All the Attributes
for (i in 4:dim(WelterweightPhy)[2])
{
  WelterweightPhy[i] = (WelterweightPhy[i]-mean(WelterweightPhy[,i]))/sd(WelterweightPhy[,i])
}

# Display the first few elements of the standarized data
head(WelterweightPhy)
```

```{r}
# Create Training and Testing Sets
num_samples = dim(WelterweightPhy)[1]
sampling.rate = 0.80
training <- sample(1:num_samples, sampling.rate * num_samples, replace=FALSE)
trainingSet <- WelterweightPhy[training, ]
testing <- setdiff(1:num_samples,training)
testingSet <- WelterweightPhy[testing, ]
```

```{r}
# Slpit Fighter's names from sets
trainingSet2 <- trainingSet[c(-1,-2)]
testingSet2 <- testingSet[c(-1,-2)]
```


# K-Nearest Neighbors

```{r}
# Get the features of the training set
trainingfeatures <- subset(trainingSet2, select=c(-Winner))
# Get the labels of the training set
traininglabels <- trainingSet2$Winner
# Get the features of the testing set
testingfeatures <- subset(testingSet2, select=c(-Winner))
```

```{r}
# Load the classification library
library(class)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(1,50, 1))
{
  predictedLabels = knn(trainingfeatures,testingfeatures,traininglabels,k=i)
sizeTestSet = dim(testingSet2)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```


```{r}
# Create Predictions Table to store predictions from different models
PredictionsTable <- data.frame(matrix(ncol = 3, nrow = dim(testingSet)[1]))
x <- c("KNN","SVM","NN")
colnames(PredictionsTable) <- x
# Fill in KNN Col
PredictionsTable$KNN <- predictedLabels
```


```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```



#SVM

```{r}
# Load the SVM Library
library(e1071)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(10,100, 10))
{
  svmModel <- svm(Winner~., data=trainingSet2, kernel="linear", cost=i)
  predictedLabels <-predict(svmModel, testingSet2)
sizeTestSet = dim(testingSet)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```

```{r}
# Fill in Table
PredictionsTable$SVM <- predictedLabels
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


# Neural Networks



```{r}
# Convert the characters to numeric values
# training
trainingSet2$Winner <- as.character(trainingSet2$Winner)
trainingSet2$Winner[trainingSet2$Winner=="Red"] = 0
trainingSet2$Winner[trainingSet2$Winner=="Blue"] = 1
trainingSet2$Winner <- as.numeric(trainingSet2$Winner)
# testing
testingSet2$Winner <- as.character(testingSet2$Winner)
testingSet2$Winner[testingSet2$Winner=="Red"] = 0
testingSet2$Winner[testingSet2$Winner=="Blue"] = 1
testingSet2$Winner <- as.numeric(testingSet2$Winner)
```

```{r}
#Load the Neural Net Library
library(neuralnet)
#Train a Neural Network. Note that the formula should be explicitly specified.
#Since we are doing a classification, we need to set linear.output=FALSE
#We will also use a Neural network with 2 hidden layers, with 7, 5 neurons respectively
nnModel <- neuralnet(Winner~., data=trainingSet2, hidden=c(7,5), linear.output=FALSE)
```

```{r}
# Plot
plot(nnModel)
```


```{r}
# Perform prdictions for the testing set
predictedLabels <- neuralnet::compute(nnModel, testingSet2[,2:7])

#The result is stored in predictedLabels$net.result
#Note that we need to round the predictions to get 0 or 1
predictedLabels<-round(predictedLabels$net.result)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(predictedLabels != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 0)
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 1)
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Fill in Table
PredictionsTable$NN <- predictedLabels
```

```{r}
# Convert #s to labels
PredictionsTable$NN <- as.character(PredictionsTable$NN)
PredictionsTable$NN[PredictionsTable$NN== "0" ] = "Red"
PredictionsTable$NN[PredictionsTable$NN== "1"] = "Blue"
```



# Convert Back
```{r}
# Convert back
# training
trainingSet2$Winner[trainingSet2$Winner== 0 ] = "Red"
trainingSet2$Winner[trainingSet2$Winner== 1 ] = "Blue"
trainingSet2$Winner <- as.factor(trainingSet2$Winner)
# testing
testingSet2$Winner[testingSet2$Winner== 0 ] = "Red"
testingSet2$Winner[testingSet2$Winner== 1 ] = "Blue"
testingSet2$Winner <- as.factor(testingSet2$Winner)
```

# Fixing class Imbalance 
Most the data has the red class winning, which should not be the case due to qualitative reasons. Both red and blue class should equally represented as possible. 

Ideally would convert winning red observations into winning blue observations, but do not know who to do that yet.

Instead, I will try to achieve this by over and under sampling. 

```{r}
library(ROSE)
```

```{r}
#See # of wins between colors
table(trainingSet$Winner)
```

```{r}
ROSE <- ovun.sample(Winner~.,data = trainingSet2, method = "both", p = 0.5, seed = 111, N = 466)$data
table(ROSE$Winner)
```



# K-NN

```{r}
# Get the features of the training set
trainingfeatures <- subset(ROSE, select=c(-Winner))
# Get the labels of the training set
traininglabels <- ROSE$Winner
# Get the features of the testing set
testingfeatures <- subset(testingSet2, select=c(-Winner))
```

```{r}
# Load the classification library
library(class)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(1,50, 1))
{
  predictedLabels = knn(trainingfeatures,testingfeatures,traininglabels,k=i)
sizeTestSet = dim(testingSet2)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Fill in Table
PredictionsTable$KNN_RB <- predictedLabels
```


#SVM Model

```{r}
# Load the SVM Library
library(e1071)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(10,100, 10))
{
  svmModel <- svm(Winner~., data=ROSE, kernel="linear", cost=i)
  predictedLabels <-predict(svmModel, testingSet2)
sizeTestSet = dim(testingSet)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Fill in Table
PredictionsTable$SVM_RB <- predictedLabels
```

# Combining predictions of models

```{r}
head(PredictionsTable)
```

```{r}
PredictionsTable$KNN <- as.character(PredictionsTable$KNN)
PredictionsTable$SVM <- as.character(PredictionsTable$SVM)
PredictionsTable$KNN_RB <- as.character(PredictionsTable$KNN_RB)
PredictionsTable$SVM_RB <- as.character(PredictionsTable$SVM_RB)
```


```{r}
PredictionsTable[PredictionsTable == "Red"] <- 1
PredictionsTable[PredictionsTable == "Blue"] <- 0
```

```{r}
PredictionsTable$KNN <- as.numeric(PredictionsTable$KNN)
PredictionsTable$SVM <- as.numeric(PredictionsTable$SVM)
PredictionsTable$NN <- as.numeric(PredictionsTable$NN)
PredictionsTable$KNN_RB <- as.numeric(PredictionsTable$KNN_RB)
PredictionsTable$SVM_RB <- as.numeric(PredictionsTable$SVM_RB)
```

```{r}
#Scoring 
PredictionsTable$Score <- rowSums(PredictionsTable)
```

```{r}
# Prediction 
PredictionsTable$Winner <- ifelse(PredictionsTable$Score > 2.5, "Red", "Blue")
```

```{r}
#Input Winner's Name
PredictionsTable$Red <- testingSet$R_fighter
PredictionsTable$Blue <- testingSet$B_fighter
PredictionsTable$Winner_Name <- ifelse(PredictionsTable$Winner == "Red", as.character(PredictionsTable$Red), as.character(PredictionsTable$Blue))
```

# Results

```{r}
ResultsM2 <- PredictionsTable
head(ResultsM2)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(PredictionsTable$Winner != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (PredictionsTable$Winner == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (PredictionsTable$Winner == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


#--------------------------------------------------------------------------------------------------------------
# Another way to fix the data imbalance is to "flip" the data from Red to Blue & COMBINE features
#--------------------------------------------------------------------------------------------------------------

#Welterweight
Read the data
```{r}
#Select Physical Features
WelterweightPhy <-  Welterweight[c(1,2,3,69,70,71,136,137,138,139,140)]
head(WelterweightPhy)
```

```{r}
WelterweightPhyFlip = WelterweightPhy[,c(3,2,4:6,10,1,7:9,11)]
WelterweightPhyFlip$B_fighter <- as.character(WelterweightPhyFlip$B_fighter)
WelterweightPhyFlip$R_fighter <- as.character(WelterweightPhyFlip$R_fighter)
head(WelterweightPhyFlip)
```

```{r}
for(i in 1:dim(WelterweightPhyFlip)[1]){
  if(WelterweightPhyFlip$Winner[i] == "Red"){
  randomNum = runif(1)
    if(randomNum < 0.25){
      tempDf = WelterweightPhyFlip[i,2:6]
      WelterweightPhyFlip[i,2:6] = WelterweightPhyFlip[i,7:11]
      WelterweightPhyFlip[i,7:11] = tempDf[1:5]
      WelterweightPhyFlip[i,1] = "Blue"
    }
  }
}
WelterweightPhyFlip$B_fighter <- as.factor(WelterweightPhyFlip$B_fighter)
WelterweightPhyFlip$R_fighter <- as.factor(WelterweightPhyFlip$R_fighter)
head(WelterweightPhyFlip)
```

```{r}
table(WelterweightPhyFlip$Winner)
WelterweightPhyFlip <- WelterweightPhyFlip[c(2,7,1,3,4,5,6,8,9,10,11)]
```

Read the data
```{r}

#Take the difference between features 
WelterweightPhyFlip$Height <- WelterweightPhyFlip$B_Height_cms -WelterweightPhyFlip$R_Height_cms
WelterweightPhyFlip$Reach <- WelterweightPhyFlip$B_Reach_cms - WelterweightPhyFlip$R_Reach_cms
WelterweightPhyFlip$Weight <- WelterweightPhyFlip$B_Weight_lbs - WelterweightPhyFlip$R_Weight_lbs
WelterweightPhyFlip$Age <- WelterweightPhyFlip$B_age - WelterweightPhyFlip$R_age
WelterweightPhyFlip$R_H_RF <- WelterweightPhyFlip$R_Reach_cms/WelterweightPhyFlip$R_Height_cms
WelterweightPhyFlip$R_H_BF <- WelterweightPhyFlip$B_Reach_cms/WelterweightPhyFlip$B_Height_cms
WelterweightPhyFlip <- WelterweightPhyFlip[c(1,2,3,12,13,14,15,16,17)]
```

# Create testing & training sets for SVM, K-NN, & Neural Networks

```{r}
# Normalize All the Attributes
for (i in 4:dim(WelterweightPhyFlip)[2])
{
  WelterweightPhyFlip[i] = (WelterweightPhyFlip[i]-mean(WelterweightPhyFlip[,i]))/sd(WelterweightPhyFlip[,i])
}

# Display the first few elements of the standarized data
head(WelterweightPhyFlip)
```

```{r}
# Create Training and Testing Sets
num_samples = dim(WelterweightPhyFlip)[1]
sampling.rate = 0.80
training <- sample(1:num_samples, sampling.rate * num_samples, replace=FALSE)
trainingSet <- WelterweightPhyFlip[training, ]
testing <- setdiff(1:num_samples,training)
testingSet <- WelterweightPhyFlip[testing, ]
```

```{r}
# Slpit Fighter's names from sets
trainingSet2 <- trainingSet[c(-1,-2)]
testingSet2 <- testingSet[c(-1,-2)]
```


# K-Nearest Neighbors

```{r}
# Get the features of the training set
trainingfeatures <- subset(trainingSet2, select=c(-Winner))
# Get the labels of the training set
traininglabels <- trainingSet2$Winner
# Get the features of the testing set
testingfeatures <- subset(testingSet2, select=c(-Winner))
```

```{r}
# Load the classification library
library(class)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(1,50, 1))
{
  predictedLabels = knn(trainingfeatures,testingfeatures,traininglabels,k=i)
sizeTestSet = dim(testingSet2)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```


```{r}
# Create Predictions Table to store predictions from different models
PredictionsTable <- data.frame(matrix(ncol = 3, nrow = dim(testingSet)[1]))
x <- c("KNN","SVM","NN")
colnames(PredictionsTable) <- x
# Fill in KNN Col
PredictionsTable$KNN <- predictedLabels
```


```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```



#SVM

```{r}
# Load the SVM Library
library(e1071)
```

```{r}
lowestRate <- 100
bestTuning <- 0
for (i in seq(10,100, 10))
{
  svmModel <- svm(Winner~., data=trainingSet2, kernel="linear", cost=i)
  predictedLabels <-predict(svmModel, testingSet2)
sizeTestSet = dim(testingSet)[1]
error = sum(predictedLabels != testingSet2$Winner)
misclassification_rate = error/sizeTestSet
if(misclassification_rate < lowestRate){
  lowestRate <- misclassification_rate
  bestTuning <- i
}
}
bestTuning
lowestRate
```

```{r}
# Fill in Table
PredictionsTable$SVM <- predictedLabels
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


# Neural Networks


```{r}
# Convert the characters to numeric values
# training
trainingSet2$Winner <- as.character(trainingSet2$Winner)
trainingSet2$Winner[trainingSet2$Winner=="Red"] = 0
trainingSet2$Winner[trainingSet2$Winner=="Blue"] = 1
trainingSet2$Winner <- as.numeric(trainingSet2$Winner)
# testing
testingSet2$Winner <- as.character(testingSet2$Winner)
testingSet2$Winner[testingSet2$Winner=="Red"] = 0
testingSet2$Winner[testingSet2$Winner=="Blue"] = 1
testingSet2$Winner <- as.numeric(testingSet2$Winner)
```

```{r}
#Load the Neural Net Library
library(neuralnet)
#Train a Neural Network. Note that the formula should be explicitly specified.
#Since we are doing a classification, we need to set linear.output=FALSE
#We will also use a Neural network with 2 hidden layers, with 7, 5 neurons respectively
nnModel <- neuralnet(Winner~., data=trainingSet2, hidden=c(7,5), linear.output=FALSE)
```

```{r}
# Plot
plot(nnModel)
```


```{r}
# Perform prdictions for the testing set
predictedLabels <- neuralnet::compute(nnModel, testingSet2[,2:7])

#The result is stored in predictedLabels$net.result
#Note that we need to round the predictions to get 0 or 1
predictedLabels<-round(predictedLabels$net.result)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(predictedLabels != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (predictedLabels == 0)
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the data points that are misclassified
IsWrong = (predictedLabels != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (predictedLabels == 1)
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Fill in Table
PredictionsTable$NN <- predictedLabels
```

```{r}
# Convert #s to labels
PredictionsTable$NN <- as.character(PredictionsTable$NN)
PredictionsTable$NN[PredictionsTable$NN== "0" ] = "Red"
PredictionsTable$NN[PredictionsTable$NN== "1"] = "Blue"
```

# Convert Back
```{r}
# Convert back
# training
trainingSet2$Winner[trainingSet2$Winner== 0 ] = "Red"
trainingSet2$Winner[trainingSet2$Winner== 1 ] = "Blue"
trainingSet2$Winner <- as.factor(trainingSet2$Winner)
# testing
testingSet2$Winner[testingSet2$Winner== 0 ] = "Red"
testingSet2$Winner[testingSet2$Winner== 1 ] = "Blue"
testingSet2$Winner <- as.factor(testingSet2$Winner)
```

# Combining predictions of models

```{r}
head(PredictionsTable)
```

```{r}
PredictionsTable$KNN <- as.character(PredictionsTable$KNN)
PredictionsTable$SVM <- as.character(PredictionsTable$SVM)
```


```{r}
PredictionsTable[PredictionsTable == "Red"] <- 1
PredictionsTable[PredictionsTable == "Blue"] <- 0
```

```{r}
PredictionsTable$KNN <- as.numeric(PredictionsTable$KNN)
PredictionsTable$SVM <- as.numeric(PredictionsTable$SVM)
PredictionsTable$NN <- as.numeric(PredictionsTable$NN)
```

```{r}
#Scoring 
PredictionsTable$Score <- rowSums(PredictionsTable)
```

```{r}
# Prediction 
PredictionsTable$Winner <- ifelse(PredictionsTable$Score > 1.5, "Red", "Blue")
```

```{r}
#Input Winner's Name
PredictionsTable$Red <- testingSet$R_fighter
PredictionsTable$Blue <- testingSet$B_fighter
PredictionsTable$Winner_Name <- ifelse(PredictionsTable$Winner == "Red", as.character(PredictionsTable$Red), as.character(PredictionsTable$Blue))
```

# Results

```{r}
ResultsFlip1 <- PredictionsTable
head(ResultsFlip1)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet2)[1]
# Get the number of data points that are misclassified
error = sum(PredictionsTable$Winner != testingSet2$Winner)
# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsR = (PredictionsTable$Winner == 'Red')
# Get the data points that are misclassified and are classified as L
IsWrongAndR = (IsWrong & IsR)
error = sum(IsWrongAndR)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```

```{r}
# Get the number of data points in the test set
sizeTestSet = dim(testingSet)[1]
# Get the data points that are misclassified
IsWrong = (PredictionsTable$Winner != testingSet2$Winner)
# Get the data points that are classified as L
IsB = (PredictionsTable$Winner == 'Blue')
# Get the data points that are misclassified and are classified as L
IsWrongAndB = (IsWrong & IsB)
error = sum(IsWrongAndB)

# Calculate the misclassification rate
misclassification_rate = error/sizeTestSet
# Display the misclassification rate
print(misclassification_rate)
```


